Studies of quantum dots

### 2016-10-18 Manuscript/paper.tex
Update spEnergiesQD/Manuscript/paper.tex

The second-quantized operators in \eqref{eq:second_quantized_one_body_hamiltonian} and \eqref{eq:second_quantized_two_body_hamiltonian} are constructed with respect to the vacuum state.  That is, the operator strings $\hat{a}_p^\dagger \hat{a}_q$ and $\hat{a}_p^\dagger \hat{a}_q^\dagger \hat{a}_s \hat{a}_r$ both have a vacuum expectation value of zero:
\begin{align*}
&\bra{\varnothing} \hat{a}_p^\dagger \hat{a}_q \ket{\varnothing} = 0 &
&\bra{\varnothing} \hat{a}_p^\dagger \hat{a}_q^\dagger \hat{a}_s \hat{a}_r \ket{\varnothing} = 0
\end{align*}
They are said to be \emph{normal-ordered} with respect to the vacuum state $\varnothing$.  It is possible to be normal-ordered with respect to a non-vacuum state, however.  Consider a Slater determinant $\Phi$ with $N_{\text F}$ occupied states:
\begin{align*}
  \ket{\Phi} \equiv \ket{\varnothing_{i_1, i_2, \ldots, i_{N_{\text F}}}}
\end{align*}
where $\{i_1, i_2, \ldots, i_{N_{\text F}}\}$ is a set of $N_{\text F}$ indices that label the occupied states.  One can define a set of \textit{quasiparticle field operators} $\hat b$ and $\hat b^\dagger$ that obey the same anticommutation relations as the $\hat a$ operators:
\begin{align*}
  \hat b_p \equiv \bar n_p \hat a_{p} + n_p \hat a_{p}^\dagger
\end{align*}
The quasiparticle field operators treat $\ket{\Phi}$ as their ``vacuum state'' (sometimes called ``Fermi vaccuum'' in an abstract sense), despite not being the physical vacuum state.  As before, it is possible to construct arbitrary Slater determinants by applying quasiparticle operators to the reference state:
\begin{align}
  \ket{\Phi_{p_1, \ldots, p_k}} \equiv \hat{b}^\dagger_{p_1} \cdots \hat{b}^\dagger_{p_k} \ket{\Phi}
  \label{eq:quasisd}
\end{align}
In nonrelativistic many-body theory, it is expected that $\{p_1, \ldots, p_k\}$ contain a balanced number of occupied and unoccupied states to conserve the total number of particles; otherwise, the Slater determinant would represent a different system.

@@

Since for this paper we are interested in the addition and removal energies, which involve one particle more or one particle less than a closed-shell state, we can take the label $u$ to be simply the label of the state that is added or removed from the closed-shell state $\Phi$.  With this realization, we can then express the perturbation expansion in terms of summations over matrix elements as we had done before for the IM-SRG flow equation, perhaps with the aid of Wick's theorem or diagrammatic techniques.  This leads to the following expression for the second-order correction:
\begin{gather*}
  \epsilon_p^{(2)} =
      \frac{1}{2} \sum_{i a b} \frac{n_i \bar{n}_a \bar{n}_b |V_{p i a b}|^2}{H^{[1]}_p + H^{[1]}_i - H^{[1]}_a - H^{[1]}_b}
    - \frac{1}{2} \sum_{i j a} \frac{n_i n_j \bar{n}_a |V_{i j p a}|^2}{H^{[1]}_i + H^{[1]}_j - H^{[1]}_p - H^{[1]}_a}
\end{gather*}
[[write out the full thing using $H^1_{p,p}$]]
The expression is depicted in Figure \ref{fig:diagrams-sfe}.  Since there are numerous terms in the third-order correction $\epsilon_p^{(3)}$, they are listed exclusively in a diagrammatic form in Figure \ref{fig:diagrams-sfe}.

@@

@@

\begin{equation}
x_{i+1} = x_i + \frac{1}{2}\delta\tau F(\mathbf{r})_x + \xi,
\end{equation}
\noindent
where $\xi$ is a uniform distributed random number with variance $\mathrm{Var}(\xi) = 2D\delta\tau$. A typical value of $N_w$ in this work is 250 000.

@@

In Figure \ref{fig:gs} we show the calculated ground state energies from
Hartree-Fock, second-order (M\o ller-Plesset) perturbation theory (MPPT), and
IMSRG.  The exact result from diffusion Monte Carlo (DMC) is also shown as a
dashed line in systems where it is known \cite{PhysRevB.84.115302}.

Both perturbation theory and IMSRG add corrections in addition to the
Hartree-Fock results to recover the correlation energy and they do quite well
in recovering the correlation energy in the system.  Compared with MPPT, IMSRG
generally converges faster, although does not always converge to the DMC
result.  Generally speaking, IMSRG is closer to the DMC result as the number
of particles increases.

There are a few cases where the IMSRG over-corrects the result, leading to a
ground state energy that is lower than the exact result.  This is not
unexpected given that, unlike Hartree-Fock, IMSRG is non-variational.  These
tend to occur when the number of particles is few, or the frequency is very
low (high correlation).  However, MPPT does not do much better in the high
frequency regimes as it often converges slower than IMSRG.

>>>>

The second-quantized operators in \eqref{eq:second_quantized_one_body_hamiltonian} and \eqref{eq:second_quantized_two_body_hamiltonian} are constructed with respect to the vacuum state.  That is, the operator strings $\hat{a}_p^\dagger \hat{a}_q$ and $\hat{a}_p^\dagger \hat{a}_q^\dagger \hat{a}_s \hat{a}_r$ both have a vacuum expectation value of zero:
\begin{align*}
&\bra{\varnothing} \hat{a}_p^\dagger \hat{a}_q \ket{\varnothing} = 0 &
&\bra{\varnothing} \hat{a}_p^\dagger \hat{a}_q^\dagger \hat{a}_s \hat{a}_r \ket{\varnothing} = 0
\end{align*}
They are said to be \emph{normal-ordered} with respect to the vacuum state $\varnothing$.  It is possible to be normal-ordered with respect to a non-vacuum state, however.  Consider a Slater determinant $\Phi$ with $N_{\text F}$ occupied states:
\begin{align*}
  \ket{\Phi} \equiv \ket{\varnothing_{i_1, i_2, \ldots, i_{N_{\text F}}}}
\end{align*}
where $\{i_1, i_2, \ldots, i_{N_{\text F}}\}$ is a set of $N_{\text F}$ indices that label the occupied states.  One can define a set of \textit{quasiparticle field operators} $\hat b$ and $\hat b^\dagger$ that obey the same anticommutation relations as the $\hat a$ operators:
\begin{align*}
  \hat b_p \equiv \bar n_p \hat a_{p} + n_p \hat a_{p}^\dagger
\end{align*}
The quasiparticle field operators treat $\ket{\Phi}$ as their ``vacuum state'' (sometimes called ``Fermi vaccuum'' in an abstract sense), despite not being the physical vacuum state.  As before, it is possible to construct arbitrary Slater determinants by applying quasiparticle operators to the reference state:
\begin{align}
  \ket{\Phi_{p_1, \ldots, p_k}} \equiv \hat{b}^\dagger_{p_1} \cdots \hat{b}^\dagger_{p_k} \ket{\Phi}
  \label{eq:quasisd}
\end{align}
If $p$ is an unoccupied state, then applying $\hat{b}_p^\dagger$ creates a particle exactly like $\hat{a}_p^\dagger$.  However, if $p$ is an occupied state, then applying $\hat{b}_p^\dagger$ destroys an existing particle in $\Phi$ or, equivalently, creates a \textit{hole} state.  The set of labels $\{p_1, \ldots, p_k\}$ may contain up to $N$ hole states and a potentially infinite number of particle states.  It is typical to classify the state $\Phi_{p_1, \ldots, p_k}$ as an $N_{\text{p}}$-particle-$N_{\text{h}}$-hole state, meaning that it contains $N_{\text{p}}$ labels that denote unoccupied states and $N_{\text{h}}$ labels that denote occupied states.  If $N_{\text{p}} = N_{\text{h}}$, the Slater determinant represents a system containing the same number of physical particles.

@@

Since for this paper we are interested in the addition and removal energies, which involve one particle more or one particle less than a closed-shell state, we can take the label $u$ to be simply the label of the state that is added or removed from the closed-shell state $\Phi$.  With this realization, we can then express the perturbation expansion in terms of summations over matrix elements as we had done before for the IM-SRG flow equation, perhaps with the aid of Wick's theorem or diagrammatic techniques.  This leads to the following expression for the second-order correction:
\begin{align*}
  \epsilon_p^{(2)}
  &=
  \frac{1}{2} \sum_{i a b} \frac{n_i \bar{n}_a \bar{n}_b |H^{[2]}_{p i a b}|^2}{H^{[1]}_p + H^{[1]}_i - H^{[1]}_a - H^{[1]}_b} \\
  &{}-
  \frac{1}{2} \sum_{i j a} \frac{n_i n_j \bar{n}_a |H^{[2]}_{i j p a}|^2}{H^{[1]}_i + H^{[1]}_j - H^{[1]}_p - H^{[1]}_a}
\end{align*}
[[write out the full thing using $H^1_{p,p}$]]
The expression is depicted in Figure \ref{fig:diagrams-sfe}.  Since there are numerous terms in the third-order correction $\epsilon_p^{(3)}$, they are listed exclusively in a diagrammatic form in Figure \ref{fig:diagrams-sfe}.

@@

\subsection{Equations of Motion Methods}

Particle attached and particle removed equations of motion (EOM) methods can be coupled with either IM-SRG or coupled cluster calculations. The principal idea is that one may construct a ladder operator $\hat{X}$ that promotes the $N$-particle ground state to any state in the $N + 1$ or $N - 1$ spectrum,
\begin{equation}\label{eq:ladder_def}
  \ket{\Psi^{(N \pm 1)}_u}  = \hat{X}^{(N \pm 1)}_u \ket{\Psi^{(N)}_0},
\end{equation}
where $\hat{X}$ is in principle a linear combination of excitation and de-excitation operators that change particle number by one,
\begin{align}
  \label{eq:gen_attached}
  \hat{X}^{(N+1)}_u &= \sum_p \bar{n}_p x_p  \hat{a}^\dagger_p + \frac{1}{2} \sum_{p, q, r} \bar{n}_p \bar{n}_q n_r x_{p, q; r} \hat{a}^\dagger_p \hat{a}^\dagger_q \hat{a}_r + \cdots,  \\
  \label{eq:gen_removed}
  \hat{X}^{(N-1)}_u &= \sum_p n_p x_p \hat{a}_p + \frac{1}{2} \sum_{p, q, r} n_p n_q \bar{n}_r  x_{p, q}{}^r \hat{a}^\dagger_r \hat{a}_q \hat{a}_p  + \cdots.
\end{align}
Substitution of \eqref{eq:ladder_def} into the energy eigenvalue problem $\hat H \ket{\Psi_u^{(N \pm 1)}} = E^{(N\pm1)}_u \ket{\Psi_u^{(N \pm 1)}}$ gives
\begin{equation}\label{eq:EOM}
  [\hat{H}, \hat{X}^{(N \pm 1)}_u] \ket{\Psi^{(N)}_0} = \epsilon_\pm \hat{X}^{(N \pm 1)}_u \ket{\Psi^{(N)}_0},
\end{equation}
which constitutes a generalized eigenvalue problem for the amplitudes $x$, where $\epsilon_\pm$ are the single-particle addition ($+$) and removal ($-$) energies. The quality of this calculation depends on the ansatz for the $N$-particle ground state, as well as the systematically improvable truncation on the ladder operators. In this work we include 1p and 2p1h excitations in the $N + 1$ ladder operator and likewise 1h and 2h1p operators for the $N - 1$ ladder operators.

\subsubsection*{EOM-IM-SRG}
After a single-reference ground state IM-SRG calculation, the Hamiltonian has been rotated such that the reference state is an eigenfunction with corresponding eigenvalue $E^{(A)}_0$, which is the correlated $A$-particle ground state energy. The EOM equation is therefore
\begin{equation}\label{eq:EOMIMSRG}
  [\bar{H},\bar{X}^{(A \pm 1)}_u] \ket{\Phi^{(A)}_0} = \omega_\pm \bar{X}^{(A \pm 1)}_u \ket{\Phi^{(A)}_0},
\end{equation}
where bars denote rotated operators. Now the reference state is used in place of the bare correlated ground state. The ground state IM-SRG procedure has
implicitly re-summed contributions from higher order excitations (3p2h, 3h2p, 4p3h,...) into the lower order amplitudes of the ladder operators (1p, 1h, 2p1h, 2h1p).

Despite these gains, the EOM calculation is still a partial diagonalization method, limited by the truncation to 2p1h and 2h1p operators. We expect $A$+1 ($A$-1) states
to be described appropriately by EOM-IM-SRG if their wavefunctions are dominated by 1p (1h) contributions in the rotated frame. We use partial norms of the EOM ladder operators
to make this determination:
\begin{align}\label{eq:partial_norms}
  n(1p) &= \sqrt{\sum_p \bar{n}_p | \bar{x}^p |^2},\\
  n(1h) &= \sqrt{\sum_p n_p | \bar{x}_p |^2}.
\end{align}
Large single particle partial norms indicate that the EOM truncation is valid for the relevant state. States with lower single particle norms should be treated by with a higher EOM approximation, which can be accomplished directly or perturbatively. {\color{red} My paper talks about this a bit more, and we have some plots. Of course it only concerns the A-body stuff}.

@@

\begin{equation}
x_{i+1} = x_i + \frac{1}{2}\delta\tau F(\mathbf{r})_x + \xi,
\end{equation}
\noindent
where $\xi$ is a uniform distributed random number with variance $\mathrm{Var}(\xi) = 2D\delta\tau$. A typical value of $N_w$ in this work is 250\,000.

@@

In Figure \ref{fig:gs} we show the calculated ground state energies from
Hartree-Fock, second-order (M\o ller-Plesset) perturbation theory (MPPT), and
IM-SRG.  The exact result from diffusion Monte Carlo (DMC) is also shown as a
dashed line in systems where it is known \cite{PhysRevB.84.115302}.

Both perturbation theory and IM-SRG add corrections in addition to the
Hartree-Fock results to recover the correlation energy and they do quite well
in recovering the correlation energy in the system.  Compared with MPPT, IM-SRG
generally converges faster, although does not always converge to the DMC
result.  Generally speaking, IM-SRG is closer to the DMC result as the number
of particles increases.

There are a few cases where the IM-SRG over-corrects the result, leading to a
ground state energy that is lower than the exact result.  This is not
unexpected given that, unlike Hartree-Fock, IM-SRG is non-variational.  These
tend to occur when the number of particles is few, or the frequency is very
low (high correlation).  However, MPPT does not do much better in the high
frequency regimes as it often converges slower than IM-SRG.
